\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2015-09-21 methodology.tex
Edits
remake in section 4

To uncover the elusive extent repair bug in Azure Storage vNext, its developers wrote a testing harness in \psharp. The developers of vNext expected that it was more likely for the bug to occur in the ExtMgr logic, rather than the EN logic. Hence, the testing harness focuses on the real ExtMgr using mocked ENs.

The testing harness consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):
%\begin{enumerate}
\begin{description}
\item[ExtentManager] acts as a thin wrapper machine for the real ExtMgr component in vNext.

\item[ExtentNode] represents a simplified mocked EN.

\item[TestingDriver] communicates with all other machines, relays messages between the \texttt{ExtentManager} and the \texttt{ExtentNode} machines, and is responsible for driving testing scenarios.

\item[RepairMonitor] collects states from all \texttt{ExtentNode} machines in the testing harness and checks desired properties, such as that an extent is replicated or repaired at the expected ENs.
\end{description}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/mocked_vnext}
\caption{Real Extent Manager with a mocked environment (each box represents one \psharp machine).}
\label{fig:azurestoremodel}
\end{figure}

@@

The real ExtMgr in vNext, which is our system-under-test, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.

\begin{figure*}[t]
\centering
\includegraphics[width=.9\linewidth]{img/extent_manager}
\caption{Internal components of the real Extent Manager in Azure Storage vNext.}
\label{fig:extentmanager}
\end{figure*}

\textbf{Internals of the real Extent Manager}.
Inside the real ExtMgr (see Figure~\ref{fig:extentmanager}), there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} data structure maintains the mapping records from extents to their hosting ENs. It is updated upon the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update extent manager's possible out-of-date view of the EN with the ground truth. \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.

@@

\textbf{Intercepting messages between machines}.
The real ExtMgr uses a network engine to send messages to the ENs. The testing harness mocks the original network engine in vNext and overrides its interface. In this way, the mocked network engine intercepts all outbound messages and relays them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. Conceivably, the mocked network engine could leverage the nondeterminism support in \psharp and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable.

%Method dispatch is the process of selecting which method, from a set of available methods with the same interface, should be invoked during a program's execution. There are two types of method dispatch: \emph{static}, which is resolved during compilation; and \emph{dynamic}, which is resolved in runtime.  \csharp (and thus \psharp) supports both static and dynamic dispatch, and provide the \texttt{virtual} modifier that can be used to declare a method which can be \emph{overridden} during runtime by an inheriting class. This capability is provided by the common language runtime (CLR) of Microsoft's .NET framework, and is a key feature of \csharp as well as other mainstream object-oriented languages.

%Using method dispatch for modeling is straightforward. The system-under-test exposes a set of APIs as \emph{virtual methods}. The developer can then \emph{override} these APIs and replace them with \emph{mocks} that will execute instead of the original implementations during systematic testing with \psharp.

%\begin{figure}[t]
%\begin{lstlisting}
%// Public interface of the real network engine
%class NetworkEngine {
  %public virtual void SendMessage(Socket s, Message msg);
  %public virtual void EnqueueMessage(Message msg);
%}
%
%// The mocked network engine used during testing
%class MockedNetEngine : NetworkEngine {
  %ExtentManager EM; // Handle to actual system-under-test
  %MachineId Env; // Handle to modeled environment
  %
  %public MockedNetEngine(ExtentManager em, MachineId env) {
    %this.EM = em;
    %this.Env = env;
  %}
  %
  %public override void SendMessage(Socket s, Message msg) {
    %PSharpRuntime.Send(this.Env, new MsgEvent(), s, msg);
  %}
  %
  %public override void EnqueueMessage(Message msg) {
    %this.EM.ProcessMessage(msg);
  %}
%}
%\end{lstlisting}
%\vspace{-2mm}
%\caption{The mocked network engine used for testing the Azure Storage vNext system.}
%\label{fig:enginecode}
%%\vspace{-2mm}
%\end{figure}

@@

// mocked engine for intercepting Extent Manager messages
class MockedNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharpRuntime.Send(this.TestingDriver,
      new MessageFromExtentManagerEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-2mm}
\caption{Mocked network engine in vNext.}
\label{fig:enginecode}
%\vspace{-2mm}
\end{figure}

%We now give an example of using dynamic dispatch to model the network engine of an extent manager in the Azure Storage vNext case study (see Figure~\ref{fig:enginecode}). The network engine is responsible for sending to and receiving messages from the various components of the system. During real execution, the network engine uses a custom remote procedure call (RPC) .NET library for communication. For testing, though, it is desirable to replace all calls to this RPC library with \psharp send and receive operations, which can be captured and systematically interleaved to find bugs. We easily achieved this by exposing the original send message operation of the network engine as a virtual method, and then overriding it for testing. In the overridden method, we created a \psharp event and then we wrapped the original message in this event's payload. Then, instead of invoking the RPC library, we invoke the \texttt{PSharpRuntime.Send(...)} method, which asynchronously sends the event (containing the original message) to the target extent node machine.

%For mocking the receive operation, we take advantage of the implicit receive of events in \psharp machines. When a extent node machine receives an event, an appropriate event handler is invoked, which extracts the original message from the payload and then handles it accordingly.

\textbf{Specifics of the ExtentManager machine}.
Figure~\ref{fig:wrap_target} shows code snippets from the \texttt{ExtentManager} \psharp machine. \texttt{ExtentManager} is a thin wrapper of the real ExtMgr. The mocked network engine replaces the real one in ExtMgr, intercepts all the outbound messages from ExtMgr and relays them to \texttt{TestingDriver}.

%A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested. The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class. The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives. This is discussed in detail in \S\ref{sec:method:model:dispath}.

%As shown in Figure~\ref{fig:wrap_target}, the \texttt{ExtentManagerMachine} is a wrapper machine and is used to connect the \psharp modeled code with the legacy \csharp code. A typical wrapper machine contains a handle (as a field) to the system component that is being wrapped (in our case the real Extent Manager object) and the event handlers are responsible for translating any received events to method calls in the real code that is being tested. The \texttt{ExtentManager} is accompanied by a mocked version of the real Network Engine class. The real Extent Manager is communicating with the outside world via remote procedure calls, but in \psharp harness we want to capture this communicating and expose it as calls to \psharp communicating primitives. This is discussed in detail in \S\ref{sec:method:model:dispath}.

\begin{figure}[t]
\begin{lstlisting}
// Wrapping the target vNext component in a P# machine
class ExtentManagerMachine : Machine {
  private ExtentManager extMgr; // real vNext code

  void Init() {
    extMgr = new ExtentManager();
    extMgr.netEngine = new MockedNetEngine(); // mock network
    extMgr.isMockingTimer = true;	 // disable internal timer
  }

  [OnEvent(MessageFromExtentNode, DeliverExtentNodeMessage)]
  void DeliverExtentNodeMessage() {
    var msg = (ExtentNodeMessage)this.Payload;
    // relay messages from Extent Node to Extent Manager
    extMgr.ProcessMessage(msg);
  }

  [OnEvent(TimerTick), nameof(ProcessExtentRepair))]
  void ProcessExtentRepair() {
    // extent repair loop driven by external timer
    extMgr.ProcessEvent(new ExtentRepairEvent());
  }
}
\end{lstlisting}
\vspace{-2mm}
\caption{System-under-test: instance of the real Extent Manager wrapped inside the \texttt{ExtentManager} machine.}
\label{fig:wrap_target}
%\vspace{-2mm}
\end{figure}

Messages coming from the ENs do \emph{not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessMessage}. The benefit of this approach is that ExtMgr can be tested without modifying its code; ExtMgr is simply unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.

System correctness should \emph{not} hinge on the frequency of any individual timer. Instead, all nondeterminism due to timing-related events must be delegated to \psharp. To achieve this, all timers inside ExtMgr are disabled, and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}. The \psharp timers send nondeterministic timeout events that are controlled by the \psharp runtime. Hence, the \psharp testing engine has the freedom to schedule arbitrary interleavings between timeout events and all other regular system events.

@@

The \texttt{ExtentNode} machine is a simplified version of the original EN. The machine omits much of the complex details of a real EN, and only mocks the logic necessary for the testing scenarios. This mocked logic includes: repairing an extent from its replica, and sending sync reports and heartbeat messages periodically to the \texttt{ExtentManager} machine.

The testing harness leverages components of Azure Storage vNext whenever it is appropriate. For example, the \texttt{ExtentNode} machine re-uses the \texttt{ExtentCenter} data structure, a component that is used inside a real EN for extent bookkeeping.

In the mocked extent repair logic, \texttt{ExtentNode} takes action upon receiving an extent repair request from the \texttt{ExtentManager} machine. It sends a copy request to a source \texttt{ExtentNode} machine where a replica is stored. After receiving an \texttt{ExtentCopyResponse} event from the source, it updates the internal \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.

In the mocked EN sync logic, the machine is again driven by an external timer provided by \psharp. It prepares a sync report with \texttt{extCtr.GetSyncReport(...)} and asynchronously sends the report to \texttt{ExtentManager} with \texttt{PSharpRuntime.Send(...)}.

%\begin{figure}[t]
%\begin{lstlisting}
%// Wrapping the target vNext component in P#
%class ExtentNodeMachine : Machine {
  %private ExtentNode.ExtentCenter extCtr; // real vNext component
%
	%[OnEventDoAction(typeof(RepairExtentRequestEvent), nameof(ProcessRepairExtentRequest))]
	%[OnEventDoAction(typeof(DownloadExtentRequestEvent), nameof(ProcessDownloadExtentRequestEvent))]
	%[OnEventDoAction(typeof(DownloadExtentResponseEvent), nameof(ProcessDownloadExtentResponseEvent))]
	%[OnEventDoAction(typeof(TimerTickEvent), nameof(ProcessExtentNodeSyncEvent))]
%
  %void ProcessRepairExtentRequest()
  %{
    %var source = GetRepairSourceMachine(this.Payload);
    %PSharpRuntime.Send(source, new DownloadExtentRequestEvent(), this.Payload);
  %}
%
  %void ProcessDownloadExtentRequestEvent()
  %{
	  %var destination = GetRepairDestinationMachine(this.Payload);
    %var ext_id = GetExtentID(this.Payload);
    %
		%Extent ext;
    %if (extCtr.TryGet(ext_id, out ext))
      %PSharpRuntime.Send(destination, new DownloadExtentResponseEvent(), ext);
  %}
%
  %void ProcessDownloadExtentResponseEvent()
  %{
    %var ext = (Extent)this.Payload;
    %extCtr.AddExtent(ext);
%
    %this.Monitor<RepairMonitor>(new NotifyExtentRepairCompletion(), this.NodeID);
  %}
%
  %void ProcessExtentNodeSyncEvent()
  %{
    %var sync = new ExtentNodeSync(this.NodeID);
    %extCtr.Extents.ForEach(ext =>
      %{
        %var rec = new ExtentRecord(ext.ExtentID, ext.Length, ext.Sealed);
        %sync.ExtentRecs.Add(rec);
      %});
    %PSharpRuntime.Send(this.NameNode, new NameNodeMessageEvent(), sync);
  %}
%}
%\end{lstlisting}
%\vspace{-2mm}
%\caption{Mocked Extent Node.}
%\label{fig:mocked_en}
%%\vspace{-2mm}
%\end{figure}

%\begin{figure}[t]
%\begin{lstlisting}
%// Mocking Extent Node in P#
%class ExtentNodeMachine : Machine {
  %// use real vNext component whenever appropriate
  %private ExtentNode.ExtentCenter extCtr;
%
  %[OnEventDoAction(typeof(RepairExtentRequestEvent), nameof(ProcessRepairExtentRequest))]
  %[OnEventDoAction(typeof(TimerTickEvent), nameof(ProcessExtentNodeSync))]
%
  %// extent repair logic
  %void ProcessRepairExtentRequest()
  %{
    %var source = GetRepairSourceMachine(this.Payload);
    %PSharpRuntime.Send(source, new DownloadExtentRequestEvent(), this.Payload);
  %}
  %...
%
  %// extent node sync logic
  %void ProcessExtentNodeSync()
  %{
    %var sync = new ExtentNodeSync(this.NodeID);
    %extCtr.Extents.ForEach(ext =>
      %{
        %sync.ExtentRecords.Add(new ExtentRecord(ext));
      %});
    %PSharpRuntime.Send(this.ExtentManagerMachine, new MessageFromExtentNodeEvent(), sync);
  %}
%}
%\end{lstlisting}
%\vspace{-2mm}
%\caption{Mocked Extent Node.}
%\label{fig:mocked_en}
%%\vspace{-2mm}
%\end{figure}

\begin{figure}[t]
\begin{lstlisting}
// Mocking Extent Node in P#
class ExtentNodeMachine : Machine {
  // leverage real vNext component whenever appropriate
  private ExtentNode.ExtentCenter extCtr;

  // extent repair logic
  ...
  [OnEvent(ExtentCopyResponse, ProcessExtentCopyResponse)]
  void ProcessExtentCopyResponse() {
    // extent copy response from source replica
    if (IsCopySucceeded(this.Payload)) {
      var rec = GetExtentRecord(this.Payload);
      extCtr.AddOrUpdate(rec); // update ExtentCenter
    }
  }

  // extent node sync logic
  [OnEvent(TimerTick, ProcessExtentNodeSync)]
  void ProcessExtentNodeSync() {
    var sync = extCtr.GetSyncReport(); // prepare sync report
    PSharpRuntime.Send(this.ExtentManagerMachine,
      new MessageFromExtentNodeEvent(), sync);
  }
}
\end{lstlisting}
\vspace{-2mm}
\caption{The mocked EN in vNext.}
\label{fig:mocked_en}
%\vspace{-2mm}
\end{figure}

@@

The \texttt{TestingDriver} machine drives two testing scenarios. In the first scenario, it launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs. It waits for the extent to be replicated at the other ENs. In the second scenario, it fails one of the \texttt{ExtentNode} machines and launches a new one. It waits for the extent to be repaired on the new EN. We omit illustrative code snippets due to space constraints.

\subsection{The RepairMonitor machine}
\label{sec:method:monitor}

The \texttt{RepairMonitor} machine is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state. Whenever an EN fails, \texttt{RepairMonitor} is notified. As soon as the number of extent replicas falls below a specified target (three replicas in the current \psharp test harness), \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired. Whenever a replica is repaired, \texttt{RepairMonitor} is notified. It transitions into the cold \emph{repaired} state, when the replica number reaches again the target, as illustrated in Figure~\ref{fig:monitor}.

@@

  // cold state: repaired
  [OnEvent(NotifyEnFailure, ProcessExtentNodeFailure)]
  cold state Repaired {
    void ProcessExtentNodeFailure() {
        var node = GetExtentNode(this.Payload);
        ExtentNode2Replica.Remove(node);
        goto Repairing;
    }
  }

  // hot state: repairing
  [OnEvent(NotifyExtentRepaired, ProcessExtentRepaired)]
  hot state Repairing {
    void ProcessExtentRepaired() {
      var node = GetExtentNode(this.Payload);
      ExtentNode2Replica[node] = true;
      if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)
        goto Repaired;
    }
  }
}
\end{lstlisting}
\vspace{-2mm}
\caption{The RepairMonitor machine.}
\label{fig:monitor}
%\vspace{-2mm}
\end{figure}

@@

%%\subsection{Modeling the environment}
%\subsection{Test Harness}
%\label{sec:method:model}
%
%The environment of a distributed system might consist of other distributed systems and services, clients, operating system timers, as well as libraries for networking or other purposes. To be able to systematically test a distributed system, this environment must be modeled and all the interactions between the environment and the system, as well as all the nondeterminism, must be captured and controlled by the \psharp runtime.
%
%%\subsubsection{Creating a test harness using \psharp}
%\subsubsection{Orchestrating Machines}
%\label{sec:method:model:harness}
%
%\begin{figure}[t]
%\centering
%\includegraphics[width=\linewidth]{img/mocked_vnext}
%\caption{The real environment of the Extent Manager is replaced with a mocked version for testing.}
%\label{fig:azurestoremodel}
%\end{figure}
%
%Before being able to use \psharp to test an existing distributed system, the environment of the components that the developer wants to test have to be modeled using the \psharp state machine APIs.  This involves creating mock classes that inherit from the \texttt{Machine} abstract class. The \texttt{Machine} class exposes methods for declaring a state machine (e.g. machine states and state transitions) as well as methods for sending events to other machines, declaring event handlers for received events, and accessing the payload from the latest received event.
%
%In the Azure Storage vNext case study, we created the following \psharp machines: \texttt{Environment}, \texttt{ExtentNode}, \texttt{ExtentManager}, and \texttt{Timer}.
%
%The \texttt{Environment} (denoted with a dashed line box in Figure~\ref{fig:azurestoremodel}) is essentially a ``god'' machine: it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute. The programmer is free to create either a finite test harness (e.g. with a predefined number of node failures) or an infinite test harness (e.g. with an unbounded number of failures), as long as any nondeterminism (e.g. when to insert a failure) is captured using the available \psharp APIs.
%
%Finally, the \texttt{Timer} machine models the operating system timer and is responsible for triggering events related to the Azure Storage vNext logic (e.g. synchronization messages, heartbeats and repair logic loops). This is discussed in detail in \S\ref{sec:method:model:timers}.
%
%\subsubsection{Modeling and injecting failures}
%\label{sec:method:model:failures}
%
%In production, each Extent Node of the Azure Storage vNext system periodically (every 5 seconds) sends a heartbeat to the Extent Manager which notifies that the Extent Node is alive. Because we want to model failures and systematically inject them using \psharp, we abstract away the heartbeat mechanism in our harness. However, the Extent Manager logic relies on time intervals to detect node failures (see Figure~\ref{fig:expiration}). The way to abstract this time-related logic and connect the real code with the modeled code is to use virtual dispatch and override the virtual \texttt{IsNodeExpired} method with a mocked version.
%
%\begin{figure}[t]
%\begin{lstlisting}
%// Real code for detecting node expiration
%public virtual bool IsNodeExpired(string id, DateTime time)
%{
  %return DateTime.Compare(time, DateTime.Now) <= 0;
%}
%
%// Mocked code for detecting node expiration
%public override bool IsNodeExpired(string id, DateTime time)
%{
  %return this.DeletedNodes.Contains(id);
%}
%\end{lstlisting}
%\vspace{-2mm}
%\caption{Abstracting the node expiration logic in the Extent Manager component of Azure Storage vNext.}
%\label{fig:expiration}
%%\vspace{-2mm}
%\end{figure}
%
%Figure~\ref{fig:expiration} presents how we mocked the node expiration detection method. Instead of comparing the time interval as in the original code, we now check if the set \texttt{DeletedNodes} contains the id of the Extent Node that we are checking for expiration. If it contains the id, then it means that the node has failed. The \texttt{Environment} machine that we have created as part of our testing \psharp harness, will nondeterministically choose a node to kill, then send the id of this killed node to the Extent Manager wrapper machine, who will in turn add it to the \texttt{DeletedNodes} set.
%
%\subsubsection{Entry point to a \psharp test harness}
%\label{sec:method:model:entrypoint}
%
%A developer can specify the entry point of a \psharp test harness similar to how unit tests are typically written. Figure~\ref{fig:entrypoint} shows the source code of the entry point for the Azure Storage vNext \psharp test harness. The programmer can declare an entry point method using the \texttt{Microsoft.PSharp.Test} attribute. When the \psharp systematic testing engine is invoked, it will scan the binary and attempt to find a static method declared with this attribute. When such a method is found (in our example the \texttt{Execute()} method), the testing engine will invoke it and start testing the program. The systematic engine can optionally run multiple iterations of the test harness, each one potentially exploring a different schedule. In each iteration, the program state will be reset (any static fields must be explicitly reset by the programmer) and the entry point will be re-invoked.
%
%\begin{figure}[t]
%\begin{lstlisting}
%[Microsoft.PSharp.Test]
%public static void Execute()
%{
  %PSharpRuntime.CreateMachine(typeof(Environment));
%}
%\end{lstlisting}
%\vspace{-2mm}
%\caption{Static \csharp method acting as the entry point to the Azure Storage vNext \psharp test harness.}
%\label{fig:entrypoint}
%%\vspace{-2mm}
%\end{figure}
%
%\subsubsection{Abstracting timers}
%\label{sec:method:model:timers}
%

>>>>

To uncover the elusive extent repair bug in Azure Storage vNext, its developers wrote a test harness using \psharp. The developers expected that it was more likely for the bug to occur in the ExtMgr logic, rather than in the EN logic. Hence, the developers focused on testing the real ExtMgr using modeled ENs. The test harness for vNext consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):
\begin{description}
\item[ExtentManager] acts as a thin wrapper machine for the real ExtMgr component in vNext (see \S\ref{sec:method:wrap_target}).

\item[ExtentNode] is a simple model of an EN (see \S\ref{sec:method:mock_en}).

\item[TestingDriver] is responsible for driving testing scenarios, relaying messages between machines, and injecting failures (see \S\ref{sec:method:driver}).

\item[Timer] exploits the nondeterministic choice generation available in \psharp to model timeouts (see \S\ref{sec:method:timer}).

\item[RepairMonitor] collects EN-related state to check if the desired liveness property is satisfied (see \S\ref{sec:method:monitor}).
\end{description}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/mocked_vnext}
\caption{Real Extent Manager with a modeled environment (each box represents one \psharp machine).}
\label{fig:azurestoremodel}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}[t]
\begin{lstlisting}
// wrapping the target vNext component in a P# machine
class ExtentManagerMachine : Machine {
  private ExtentManager ExtMgr; // real vNext code

  void Init() {
    ExtMgr = new ExtentManager();
    ExtMgr.netEngine = new MockedNetEngine(); // mock network
    ExtMgr.isMockingTimer = true;	 // disable internal timer
  }

  [OnEvent(ENMsgEvent, DeliverExtentNodeMessage)]
  void DeliverExtentNodeMessage() {
    var msg = (ExtentNodeMessage)this.Payload;
    // relay messages from Extent Node to Extent Manager
    ExtMgr.ProcessMessage(msg);
  }

  [OnEvent(TimerTickEvent, ProcessExtentRepair)]
  void ProcessExtentRepair() {
    // extent repair loop driven by external timer
    ExtMgr.ProcessEvent(new ExtentRepairEvent());
  }
}
\end{lstlisting}
\vspace{-3mm}
\caption{The real Extent Manager is wrapped inside the \texttt{ExtentManager} \psharp machine.}
\label{fig:wrap_target}
%\vspace{-2mm}
\end{figure}

The real ExtMgr in vNext, which is our system-under-test, is wrapped inside the \texttt{ExtentManager} machine, as illustrated in the code snippet of Figure~\ref{fig:wrap_target}.

\begin{figure*}[t]
\centering
\includegraphics[width=.9\linewidth]{img/extent_manager}
\caption{Internal components of the real Extent Manager in Azure Storage vNext.}
\label{fig:extentmanager}
\vspace{-3mm}
\end{figure*}

\textbf{Internals of the real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} data structure maintains the mapping records from extents to their hosting ENs. It is updated upon the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update ExtMgr's possible out-of-date view of the EN with the ground truth. \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.

@@

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to the ENs. The \psharp test harness mocks the original network engine in vNext by overriding its interface. This enables the mocked network engine to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.

@@

// mocked engine for intercepting Extent Manager messages
class MockedNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharpRuntime.Send(TestingDriver,
      new ExtentManagerMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-3mm}
\caption{Mocked network engine in vNext.}
\label{fig:enginecode}
\vspace{-2mm}
\end{figure}

Intercepting all network messages and dispatching them through the \psharp runtime is important for two reasons. First, it allows \psharp to systematically explore the interleavings between asynchronous event handlers in the system. Second, the mocked network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable (not shown in this example).

Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}). The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it running in a real distributed environment and communicating with real ENs.

@@

\begin{figure}[t]
\begin{lstlisting}
// modeling Extent Node in P#
class ExtentNodeMachine : Machine {
  // leverage real vNext component whenever appropriate
  private ExtentNode.ExtentCenter ExtCtr;

  // extent repair logic
  ...
  [OnEvent(ExtentCopyRespEvent, ProcessExtentCopyResponse)]
  void ProcessExtentCopyResponse() {
    // extent copy response from source replica
    if (IsCopySucceeded(this.Payload)) {
      var rec = GetExtentRecord(this.Payload);
      ExtCtr.AddOrUpdate(rec); // update ExtentCenter
    }
  }

  // extent node sync logic
  [OnEvent(TimerTickEvent, ProcessExtentNodeSync)]
  void ProcessExtentNodeSync() {
    var sync = ExtCtr.GetSyncReport(); // prepare sync report
    this.Send(ExtentManagerMachine, new ENMsgEvent(), sync);
  }

  // extent node failure logic
  [OnEvent(FailEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the Extent Manager that this EN failed
    this.Send(ExtentManagerMachine, new ENMsgEvent(),
      new ExtentNodeDeletion(ExtentId));
    // notifies the monitor that this EN failed
    this.Monitor<RepairMonitor>(new ENFailedEvent(), this);
    this.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-3mm}
\caption{The modeled EN in vNext.}
\label{fig:mocked_en}
\vspace{-2mm}
\end{figure}

The \texttt{ExtentNode} machine is a simplified version of the original EN. The machine omits most of the complex details of a real EN, and only models the necessary logic for testing. This modeled logic includes: repairing an extent from its replica, and sending sync reports and heartbeat messages periodically to \texttt{ExtentManager}.

The \psharp test harness leverages components of the real vNext system whenever it is appropriate. For example, \texttt{ExtentNode} re-uses the \texttt{ExtentCenter} data structure, which is used inside a real EN for extent bookkeeping. In the modeled extent repair logic, \texttt{ExtentNode} takes action upon receiving an extent repair request from the \texttt{ExtentManager} machine. It sends a copy request to a source \texttt{ExtentNode} machine where a replica is stored. After receiving an \texttt{ExtentCopyRespEvent} event from the source machine, it updates the \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.

In the modeled EN sync logic, the machine is driven by an external timer modeled in \psharp (see \S\ref{sec:method:timer}). It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharpRuntime.Send(...)}. The \texttt{ExtentNode} machine also includes failure-related logic, which is discussed in detail in \S\ref{sec:method:driver}.

@@

\begin{figure}[t]
\begin{lstlisting}
// machine for driving testing scenarios in vNext
class TestingDriver : Machine {
  private HashSet<MachineId> ExtentNodes; // EN machines

  // other test harness driving logic
  ...
  void InjectNodeFailure() {
    // nondeterministically choose an EN using P#
    var node = (MachineId)this.Nondet(ExtentNodes);    
    this.Send(node, new FailEvent()); // fail the chosen EN
  }
}
\end{lstlisting}
\vspace{-3mm}
\caption{The TestingDriver machine in vNext.}
\label{fig:testing_driver}
\vspace{-2mm}
\end{figure}

The \texttt{TestingDriver} machine drives two testing scenarios. In the first scenario, \texttt{TestingDriver} launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs. \texttt{TestingDriver} then waits for the extent to be replicated at the remaining ENs. In the second testing scenario, \texttt{TestingDriver} fails one of the \texttt{ExtentNode} machines and launches a new one. \texttt{TestingDriver} then waits for the extent to be repaired on the newly launched \texttt{ExtentNode} machine.

Figure~\ref{fig:testing_driver} illustrates how \psharp can be used to inject a nondeterministic failure. \texttt{TestingDriver} can call the \texttt{InjectNodeFailure} method, which invokes the \psharp method \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of available \texttt{ExtentNode} machines. \texttt{TestingDriver} then sends the chosen machine a \texttt{FailEvent} event. The chosen \texttt{ExtentNode} handles this event (see Figure~\ref{fig:mocked_en}) by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sends it to the \texttt{ExtentManager} machine. \texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for unit testing, which maximizes code re-use. Next, the chosen \texttt{ExtentNode} notifies the liveness monitor that an EN has failed. Finally, the chosen \texttt{ExtentNode} calls the \psharp method \texttt{Halt()}, which causes the machine to terminate. \psharp is able to not only enumerate interleavings of asynchronous event handlers, but also calls to \texttt{Nondet()}, and thus failure injections.

\subsection{The Timer machine}
\label{sec:method:timer}

\begin{figure}[t]
\begin{lstlisting}
// modeling timer expiration in P#
class Timer : Machine {
  MachineId Target; // target machine

  [OnEvent(RepeatEvent, GenerateTimeTick)]
  void GenerateTimeTick() {
    // nondeterministic choice controlled by P#
    if (this.Nondet())
      this.Send(this.Target, new TimerTickEvent());
    this.Send(this, new RepeatEvent()); // loop
  }
}
\end{lstlisting}
\vspace{-3mm}
\caption{Modeling timer expiration using \psharp.}
\label{fig:mocked_timer}
\vspace{-2mm}
\end{figure}

System correctness should \emph{not} hinge on the frequency of any individual timer. Instead, all nondeterminism due to timing-related events must be delegated to \psharp. To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}. Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.

Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine of vNext. \texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice that is controlled by the \psharp runtime. Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to a target machine. The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.

\subsection{The RepairMonitor liveness monitor}
\label{sec:method:monitor}

\texttt{RepairMonitor} is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state. Whenever an EN fails, the monitor is notified with an \texttt{ENFailedEvent} event. As soon as the number of extent replicas falls below a specified target (three replicas in the current \psharp test harness), the monitor transitions into the hot \emph{repairing} state, waiting for all missing replicas to be repaired. Whenever an extent replica is repaired, \texttt{RepairMonitor} is notified with an \texttt{ExtentRepairedEvent} event. When the replica number reaches again the target, the monitor transitions into the cold \emph{repaired} state, as illustrated in the code snippet of Figure~\ref{fig:monitor}.

@@

  // cold state: repaired
  [OnEvent(ENFailedEvent, ProcessExtentNodeFailure)]
  cold state Repaired {
    void ProcessExtentNodeFailure() {
        var node = GetExtentNode(this.Payload);
        ExtentNode2Replica.Remove(node);
        goto Repairing;
    }
  }

  // hot state: repairing
  [OnEvent(ExtentRepairedEvent, ProcessExtentRepaired)]
  hot state Repairing {
    void ProcessExtentRepaired() {
      var node = GetExtentNode(this.Payload);
      ExtentNode2Replica[node] = true;
      if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)
        goto Repaired;
    }
  }
}
\end{lstlisting}
\vspace{-3mm}
\caption{The RepairMonitor liveness monitor.}
\label{fig:monitor}
\vspace{-2mm}
\end{figure}

@@
